---
title: "activerse"
output:
  github_document:
    toc: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  message = FALSE,
  warning = FALSE
)
```

<!-- badges: start -->
[![R-CMD-check](https://github.com/jhuwit/activerse/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/jhuwit/activerse/actions/workflows/R-CMD-check.yaml)
<!-- badges: end -->
  
  
<!-- activerse-intro:start -->
`activerse` is a lightweight helper package that attaches [`actibase`](https://github.com/jhuwit/actibase), [`actiread`](https://github.com/jhuwit/actiread), [`actimetrics`](https://github.com/jhuwit/actimetrics), [`actisensorlog`](https://github.com/jhuwit/actisensorlog), [`actisleep`](https://github.com/jhuwit/actisleep), [`actiwalkability`](https://github.com/jhuwit/actiwalkability), and [`actiquantiles`](https://github.com/jhuwit/actiquantiles) together.
<!-- activerse-intro:end -->


It is designed to behave a bit like `tidyverse`: load one package, then use the
functions from the actigraphy packages directly.

## Installation

```{r, eval=FALSE}
# install.packages("remotes")
remotes::install_github("jhuwit/activerse")
```

## Packages

<!-- activerse-packages:start -->
| Package | Repository | Version / status |
| --- | --- | --- |
| `actibase` | [jhuwit/actibase](https://github.com/jhuwit/actibase) | [![CRAN version](https://www.r-pkg.org/badges/version/actibase)](https://cran.r-project.org/package=actibase) [![GitHub version](https://img.shields.io/github/r-package/v/jhuwit/actibase?label=GitHub)](https://github.com/jhuwit/actibase) [![R CMD check](https://github.com/jhuwit/actibase/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/jhuwit/actibase/actions/workflows/R-CMD-check.yaml) |
| `actiread` | [jhuwit/actiread](https://github.com/jhuwit/actiread) | [![CRAN version](https://www.r-pkg.org/badges/version/actiread)](https://cran.r-project.org/package=actiread) [![GitHub version](https://img.shields.io/github/r-package/v/jhuwit/actiread?label=GitHub)](https://github.com/jhuwit/actiread) [![R CMD check](https://github.com/jhuwit/actiread/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/jhuwit/actiread/actions/workflows/R-CMD-check.yaml) |
| `actimetrics` | [jhuwit/actimetrics](https://github.com/jhuwit/actimetrics) | [![GitHub version](https://img.shields.io/github/r-package/v/jhuwit/actimetrics?label=GitHub)](https://github.com/jhuwit/actimetrics) [![R CMD check](https://github.com/jhuwit/actimetrics/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/jhuwit/actimetrics/actions/workflows/R-CMD-check.yaml) |
| `actisensorlog` | [jhuwit/actisensorlog](https://github.com/jhuwit/actisensorlog) | [![GitHub version](https://img.shields.io/github/r-package/v/jhuwit/actisensorlog?label=GitHub)](https://github.com/jhuwit/actisensorlog) [![R CMD check](https://github.com/jhuwit/actisensorlog/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/jhuwit/actisensorlog/actions/workflows/R-CMD-check.yaml) |
| `actiwalkability` | [jhuwit/actiwalkability](https://github.com/jhuwit/actiwalkability) | [![GitHub version](https://img.shields.io/github/r-package/v/jhuwit/actiwalkability?label=GitHub)](https://github.com/jhuwit/actiwalkability) [![R CMD check](https://github.com/jhuwit/actiwalkability/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/jhuwit/actiwalkability/actions/workflows/R-CMD-check.yaml) |
| `actiquantiles` | [jhuwit/actiquantiles](https://github.com/jhuwit/actiquantiles) | [![GitHub version](https://img.shields.io/github/r-package/v/jhuwit/actiquantiles?label=GitHub)](https://github.com/jhuwit/actiquantiles) [![R CMD check](https://github.com/jhuwit/actiquantiles/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/jhuwit/actiquantiles/actions/workflows/R-CMD-check.yaml) |
| `actisleep` | [jhuwit/actisleep](https://github.com/jhuwit/actisleep) | [![GitHub version](https://img.shields.io/github/r-package/v/jhuwit/actisleep?label=GitHub)](https://github.com/jhuwit/actisleep) [![R CMD check](https://github.com/jhuwit/actisleep/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/jhuwit/actisleep/actions/workflows/R-CMD-check.yaml) |

## Dependencies

The following packages support the activerse but are not themselves activerse
packages.

| Package | Repository | Version / status |
| --- | --- | --- |
| `actinet` | [jhuwit/actinet](https://github.com/jhuwit/actinet) | [![CRAN version](https://www.r-pkg.org/badges/version/actinet)](https://cran.r-project.org/package=actinet) [![GitHub version](https://img.shields.io/github/r-package/v/jhuwit/actinet?label=GitHub)](https://github.com/jhuwit/actinet) [![R CMD check](https://github.com/jhuwit/actinet/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/jhuwit/actinet/actions/workflows/R-CMD-check.yaml) |
| `asleep` | [jhuwit/asleep](https://github.com/jhuwit/asleep) | [![CRAN version](https://www.r-pkg.org/badges/version/asleep)](https://cran.r-project.org/package=asleep) [![GitHub version](https://img.shields.io/github/r-package/v/jhuwit/asleep?label=GitHub)](https://github.com/jhuwit/asleep) [![R CMD check](https://github.com/jhuwit/asleep/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/jhuwit/asleep/actions/workflows/R-CMD-check.yaml) |
| `mapnhanespa` | [jhuwit/mapnhanespa](https://github.com/jhuwit/mapnhanespa) | [![CRAN version](https://www.r-pkg.org/badges/version/mapnhanespa)](https://cran.r-project.org/package=mapnhanespa) [![GitHub version](https://img.shields.io/github/r-package/v/jhuwit/mapnhanespa?label=GitHub)](https://github.com/jhuwit/mapnhanespa) [![R CMD check](https://github.com/jhuwit/mapnhanespa/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/jhuwit/mapnhanespa/actions/workflows/R-CMD-check.yaml) |
| `sleeper` | [jhuwit/sleeper](https://github.com/jhuwit/sleeper) | [![GitHub version](https://img.shields.io/github/r-package/v/jhuwit/sleeper?label=GitHub)](https://github.com/jhuwit/sleeper) [![R CMD check](https://github.com/jhuwit/sleeper/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/jhuwit/sleeper/actions/workflows/R-CMD-check.yaml) |
| `stepcount` | [jhuwit/stepcount](https://github.com/jhuwit/stepcount) | [![CRAN version](https://www.r-pkg.org/badges/version/stepcount)](https://cran.r-project.org/package=stepcount) [![GitHub version](https://img.shields.io/github/r-package/v/jhuwit/stepcount?label=GitHub)](https://github.com/jhuwit/stepcount) [![R CMD check](https://github.com/jhuwit/stepcount/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/jhuwit/stepcount/actions/workflows/R-CMD-check.yaml) |
| `walking` | [jhuwit/walking](https://github.com/jhuwit/walking) | [![CRAN version](https://www.r-pkg.org/badges/version/walking)](https://cran.r-project.org/package=walking) [![GitHub version](https://img.shields.io/github/r-package/v/jhuwit/walking?label=GitHub)](https://github.com/jhuwit/walking) [![R CMD check](https://github.com/jhuwit/walking/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/jhuwit/walking/actions/workflows/R-CMD-check.yaml) |
<!-- activerse-packages:end -->


# `activerse` Norms and Expectations

One of the goals of the `activerse` is standardization and modularization of code. We aim for some norms and expectation for integration of new packages into the `activerse` ecosystem.  These are not hard rules, but rather guidelines for consistency and maintainability.

1. When possible, user-exposed (e.g. in `exports` in NAMESPACE) functions should have the prefix `acti_`.
2. Optional arguments are initialized in the function using `NULL` and/or use a `match.arg` case when options are presented.
3. Functions should use `actibase` helpers where they can, especially `get_sample_rate` and `get_dynamic_range`.  In cases where `sample_rate` can be passed as well as data, `sample_rate` will be initialized as `NULL` or `get_sample_rate(data)`.
4. Python dependencies are in `R/zzz.R` and wrapped using `reticulate::py_require` in the `.onLoad` function.  The `.onLoad` should call an exported `py_require_PACKAGE` where `PACKAGE` is usually the package or python module.
5. For functions that call python modules, it is encouraged for a `callr::r` wrapper to be included (this cause a `callr` dependency, so it can be Suggests.  These wrappers should have the prefix `py_acti`, for example see `actimetrics::py_acti_calculate_stepcount`, which calls `actimetrics::acti_calculate_stepcount`.  
6. Packages have vignettes where possible.  They can use files from `activerse` packages using `system.file` if they want to use the data.  
7. Individual packages can be depended on with their own system, but functions should be wrapped using the above norms into an `activerse` package to be "included".
8. Use `Roxygen: list(markdown = TRUE)` in `DESCRIPTION`, use the `usethis` package to write your package and `devtools::release()` to submit to CRAN.



# Usage

## Load the packages

```{r}
library(activerse)

# activerse-load:start
search()[grepl("^package:(actibase|actiread|actimetrics|actisensorlog|actisleep|actiwalkability|actiquantiles)$", search())]
# activerse-load:end
```

## Example data

The packages ship with small example datasets that are useful for quick tests.

```{r}
data(acti_raw_data, package = "actibase")
data(acti_count_data, package = "actimetrics")

head(acti_raw_data)
head(acti_count_data)
```

## Use `actimetrics`

Once `activerse` is attached, functions from the activerse packages are
available directly. You can also call them with the relevant `pkg::` prefix if
you prefer being explicit.

### Calculate ENMO

```{r}
enmo <- acti_calculate_enmo(data = acti_raw_data)
head(enmo)
```

### Calculate activity index

```{r}
activity_index <- acti_calculate_activity_index(
  acti_raw_data,
  unit = "1 min",
  ensure_all_time = TRUE,
  verbose = FALSE
)

head(activity_index)
```

### Calculate MAD

```{r}
mad <- acti_calculate_mad(data = acti_raw_data)
head(mad)
```

## More examples

`actimetrics` also includes functions for wear/non-wear detection, step count
estimation, and other processing helpers:

```{r, eval=FALSE}
acti_process(acti_raw_data)
acti_calculate_wear(acti_raw_data)
acti_calculate_nonwear(acti_raw_data)
acti_calculate_stepcount(acti_raw_data)
```

## Running Walking with Different Python Environments

If you want to use both walking estimation from `forest` and `stepcount`, you can run them in separate R processes using `callr` to avoid Python package conflicts. 
Here's how you can do it:

```{r, cache = TRUE}
library(callr)
library(activerse)
gt3x_path = acti_example_gt3x()
data <- acti_read_gt3x(gt3x_path, cleanup = TRUE)

stepcount_callr = function(data,
                           ...) {
  
  reticulate::py_require("stepcount==3.11.0", python_version = "3.10")
  sc <- reticulate::import("stepcount")
  stepcount::stepcount_check()
  
  res = actimetrics::acti_calculate_stepcount(data, ...)
  return(res)
}

# 2. Run the isolated background R process
result <- callr::r(
  func = stepcount_callr,
  show = TRUE,
  args = list(data = data) # Safely injects data into the process
)
head(result)

forest_callr = function(data,
                        ...) {
  reticulate::py_require(
    "git+https://github.com/onnela-lab/forest@45fb41038bd46c25d9e6a4442aa74fa03b501317", 
    python_version = "3.11")
  fr = reticulate::import("forest")
  oak = fr$oak$base
  oak
  res = actimetrics::acti_calculate_forest(data, ...)
  return(res)
}

# 2. Run the isolated background R process
fresult <- callr::r(
  func = forest_callr,
  show = TRUE,
  args = list(data = data) # Safely injects data into the process
)
head(fresult)
```


In the previous example we use `reticulate::py_require` to install python dependencies on demand, but recommend an environment (e.g. conda) with the required packages pre-installed for better performance and use on multiple subjects. 

## Updating activerse

When a new package is added to `activerse`, update the package list, the README,
and the attach helper by running:

```{r, eval=FALSE}
./scripts/update-activerse.R actisensorlog
```

If the new package lives under a different GitHub owner, pass `--owner=`:

```{r, eval=FALSE}
./scripts/update-activerse.R --owner=some-owner actisensorlog
```

The script updates `DESCRIPTION`, `R/activerse.R`, and both README files in one
pass.
