---
title: "activerse"
output:
  github_document:
    toc: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  message = FALSE,
  warning = FALSE
)
```

`activerse` is a lightweight helper package that attaches `actibase`, `actiread`,
and `actimetrics` together.

It is designed to behave a bit like `tidyverse`: load one package, then use the
functions from the actigraphy packages directly.

## Installation

```{r, eval=FALSE}
# install.packages("remotes")
remotes::install_github("jhuwit/activerse")
```

## Load the packages

```{r}
library(activerse)

search()[grepl("^package:(actibase|actiread|actimetrics)$", search())]
```

## Example data

The packages ship with small example datasets that are useful for quick tests.

```{r}
data(acti_raw_data, package = "actibase")
data(acti_count_data, package = "actimetrics")

acti_raw_data
acti_count_data
```

## Use `actimetrics`

Once `activerse` is attached, `actimetrics` functions are available directly.
You can also call them with the `actimetrics::` prefix if you prefer being
explicit.

### Calculate counts

```{r}
counts <- acti_calculate_counts(acti_raw_data, verbose = FALSE)
head(counts)
```

### Calculate activity index

```{r}
activity_index <- acti_calculate_activity_index(
  acti_raw_data,
  unit = "1 min",
  ensure_all_time = TRUE,
  verbose = FALSE
)

head(activity_index)
```

### Calculate MIMS

```{r}
mims <- acti_calculate_mims(
  acti_raw_data,
  unit = "1 min",
  ensure_all_time = TRUE
)

head(mims)
```

## More examples

`actimetrics` also includes functions for wear/non-wear detection, step count
estimation, and other processing helpers:

```{r, eval=FALSE}
acti_process(acti_raw_data)
acti_calculate_wear(acti_raw_data)
acti_calculate_nonwear(acti_raw_data)
acti_calculate_stepcount(acti_raw_data)
```

