---
title: "activerse"
output:
  github_document:
    toc: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  message = FALSE,
  warning = FALSE
)
```

`activerse` is a lightweight helper package that attaches
[`actibase`](https://github.com/jhuwit/actibase),
[`actiread`](https://github.com/jhuwit/actiread), and
[`actimetrics`](https://github.com/jhuwit/actimetrics) together.

It is designed to behave a bit like `tidyverse`: load one package, then use the
functions from the actigraphy packages directly.

## Installation

```{r, eval=FALSE}
# install.packages("remotes")
remotes::install_github("jhuwit/activerse")
```

## Packages

| Package | Repository | R CMD check |
| --- | --- | --- |
| `actibase` | [jhuwit/actibase](https://github.com/jhuwit/actibase) | [![R CMD check](https://github.com/jhuwit/actibase/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/jhuwit/actibase/actions/workflows/R-CMD-check.yaml) |
| `actiread` | [jhuwit/actiread](https://github.com/jhuwit/actiread) | [![R CMD check](https://github.com/jhuwit/actiread/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/jhuwit/actiread/actions/workflows/R-CMD-check.yaml) |
| `actimetrics` | [jhuwit/actimetrics](https://github.com/jhuwit/actimetrics) | [![R CMD check](https://github.com/jhuwit/actimetrics/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/jhuwit/actimetrics/actions/workflows/R-CMD-check.yaml) |

## Load the packages

```{r}
library(activerse)

search()[grepl("^package:(actibase|actiread|actimetrics)$", search())]
```

## Example data

The packages ship with small example datasets that are useful for quick tests.

```{r}
data(acti_raw_data, package = "actibase")
data(acti_count_data, package = "actimetrics")

head(acti_raw_data)
head(acti_count_data)
```

## Use `actimetrics`

Once `activerse` is attached, `actimetrics` functions are available directly.
You can also call them with the `actimetrics::` prefix if you prefer being
explicit.

### Calculate ENMO

```{r}
enmo <- acti_calculate_enmo(data = acti_raw_data)
head(enmo)
```

### Calculate activity index

```{r}
activity_index <- acti_calculate_activity_index(
  acti_raw_data,
  unit = "1 min",
  ensure_all_time = TRUE,
  verbose = FALSE
)

head(activity_index)
```

### Calculate MAD

```{r}
mad <- acti_calculate_mad(data = acti_raw_data)
head(mad)
```

## More examples

`actimetrics` also includes functions for wear/non-wear detection, step count
estimation, and other processing helpers:

```{r, eval=FALSE}
acti_process(acti_raw_data)
acti_calculate_wear(acti_raw_data)
acti_calculate_nonwear(acti_raw_data)
acti_calculate_stepcount(acti_raw_data)
```
