---
title: "Processing Physical Activity to a Normalized Quantile"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Processing Physical Activity to a Normalized Quantile}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

This report follows the structure of the original project page at
[johnmuschelli.com/upper_limb_gt3x_prosthesis](https://johnmuschelli.com/upper_limb_gt3x_prosthesis/)
and the source document at
[muschellij2/upper_limb_gt3x_prosthesis/index.Rmd](https://github.com/muschellij2/upper_limb_gt3x_prosthesis/blob/master/index.Rmd).
The goal is the same: show a fully reproducible pipeline for going from raw
GT3X+ data to a summary that can be mapped with participant demographics.

The report is written so that every download is stored on disk under `data/`
instead of `tempfile()`. That keeps the workflow repeatable across knits and
prevents the slow network work from running over and over.

## Setup

We first load the packages used by the analysis and tell `reticulate` which
Python environment to manage. The Python `stepcount` module is imported
explicitly so the step-counting part of the pipeline is reproducible too.

```{r setup}
has_reticulate = requireNamespace("reticulate", quietly = TRUE)
if (has_reticulate) {
  Sys.setenv(
    RETICULATE_PYTHON = "managed",
    PATH = paste(path.expand("~/.local/bin"), Sys.getenv("PATH"), sep = ":"),
    # R_USER_CACHE_DIR = file.path("data", "r-cache"),
    UV_CACHE_DIR = file.path("data", "uv-cache")
    # UV_PYTHON_PREFERENCE = "only-system",
    # UV_OFFLINE = "1",
    # UV_PRERELEASE = "allow"
  )
  library(reticulate)
  reticulate::py_require("stepcount==3.11.0", python_version = "3.10")
  sc <- reticulate::import("stepcount")
  sc$stepcount
  if (requireNamespace("stepcount", quietly = TRUE)) {
    stepcount::stepcount_check()
  }
}

library(dplyr)
library(tidyr)
library(janitor)
if (requireNamespace("readxl", quietly = TRUE)) {
  library(readxl)
}
library(activerse)
```

The report keeps downloaded files in a small directory structure so that each
artifact only has to be fetched once.

```{r}
raw_dir <- file.path("data", "raw")
derived_dir <- file.path("data", "derived")
dir.create(raw_dir, recursive = TRUE, showWarnings = FALSE)
dir.create(derived_dir, recursive = TRUE, showWarnings = FALSE)

download_if_missing <- function(url, dest) {
  dir.create(dirname(dest), recursive = TRUE, showWarnings = FALSE)
  if (!file.exists(dest)) {
    download.file(url, destfile = dest, quiet = FALSE, mode = "wb")
  }
  dest
}
```

## Goal

The data come from a publicly released study of upper-limb activity in people
with and without upper-limb prostheses. The original project pairs raw GT3X+
files with a metadata spreadsheet, then uses those data to build an analysis
that can be checked end-to-end with open-source tooling.

This report keeps the same logic, but it emphasizes why each step exists:

1. Download and cache the raw accelerometer file.
2. Read the file with an open-source GT3X reader.
3. Process the raw samples into minute-level and daily summaries.
4. Rebuild the demographic table from the spreadsheet used in
`demog_mapping.R`.
5. Combine the daily activity summary with the demographics and generate the
mapping display.

## Download Raw GT3X+

The sample GT3X+ file is downloaded once and kept on disk. That makes the
report self-contained after the first run and avoids repeated network access.

```{r acti_gt3x}
gt3x_url <- "https://ndownloader.figshare.com/files/21855555"
gt3x_path <- download_if_missing(
  gt3x_url,
  file.path(raw_dir, "AI1_NEO1B41100255_2016-10-17.gt3x.gz")
)
gt3x_path
```

## Read the File

The raw file is read directly from the gzipped GT3X+ archive. No temporary copy
is needed because the reader can work with the file in place.

```{r acti_read}
data <- acti_read_gt3x(gt3x_path, cleanup = TRUE)
data
```

## Process the Accelerometry

The helper below mirrors the original script, but the comments explain the
purpose of each stage. Optional calibration is kept as an argument so the same
function can be reused if a later analysis needs it.

```{r acti_run_code}
checker = function(data) {
  data = data %>%
    dplyr::mutate(
      date = as_date_safe(time),
      minute = lubridate::floor_date(time, unit = "minutes"),
      hourtime = hms::as_hms(time))
  
  data = data %>%
    dplyr::distinct(date, minute, hourtime, wear)
  
  # Check for duplicates - should not be there
  suppressMessages({
    dupes = janitor::get_dupes(data, date, minute)
  })
  if (nrow(dupes) > 0) {
    stop("There are duplicate rows of date and minute!")
  }
}
acti_run <- function(data, calibrate = FALSE,
                     run_stepcount = TRUE,
                     run_forest = FALSE) {
  if (calibrate) {
    data <- acti_calibrate(data)
  }
  
  message("Processing Data")
  # The reader already returns resampled data, so we can start processing.
  processed <- acti_process(data)
  
  if (run_stepcount) {
    if (requireNamespace("stepcount", quietly = TRUE) ||
        !stepcount::have_stepcount()) {
      run_stepcount = FALSE
      cli::cli_warn(paste0("`run_stepcount` = TRUE, but 
                    {.help [have_stepcount](stepcount::have_stepcount)} ", 
                    "came back `FALSE`"))
    }
  }
  if (run_stepcount) {
    message("Running Stepcount")
    # Step counts come from the Python-backed stepcount implementation.
    steps <- acti_calculate_stepcount(data) %>% 
      rename(steps_stepcount_ssl = steps) 
    
    processed <- processed %>%
      dplyr::full_join(steps)
    
    steps <- acti_calculate_stepcount(data, model_type = "rf") %>% 
      rename(steps_stepcount_rf = steps) %>% 
      select(-any_of("walking"))
    processed <- processed %>%
      dplyr::full_join(steps)
  }
  
  # Verisense steps  
  steps <- acti_calculate_verisense(data, resample_to_15hz = TRUE,
                                    method = "original") %>% 
    rename(steps_vs_original = steps)
  processed <- processed %>%
    dplyr::full_join(steps)
  
  # Verisense steps Revised
  steps <- acti_calculate_verisense(data, resample_to_15hz = TRUE,
                                    method = "revised") %>% 
    rename(steps_vs_revised = steps)
  processed <- processed %>%
    dplyr::full_join(steps)
  
  steps <- acti_calculate_sdt(data) %>% 
    rename(steps_sdt = steps)
  processed <- processed %>%
    dplyr::full_join(steps)
  
  if (run_forest) {
    if (!walking::have_forest()) {
      run_forest = FALSE
      cli::cli_warn(paste0("`run_forest` = TRUE, but 
                    {.help [have_forest](walking::have_forest)} ", 
                    "came back `FALSE`"))
    }
  }
  if (run_forest) {  
    steps <- acti_calculate_forest(data) %>% 
      rename(steps_forest = steps)
    processed <- processed %>%
      dplyr::full_join(steps)  
  }
  
  
  
  # Flag which days have enough valid wear time to count as included.
  inclusion <- create_day_inclusion(processed, min_required = 1368L)
  processed <- processed %>%
    mutate(date = as.Date(time)) %>%
    left_join(inclusion)
  
  # A small helper for daily totals that ignore missing values.
  na_sum <- function(x) sum(x, na.rm = TRUE)
  
  # Collapse the minute-level data to daily totals for the measures we care about.
  daily <- processed %>%
    acti_create_date() %>%
    group_by(date) %>%
    summarise(
      dplyr::across(
        c(any_of(c("counts", "counts_log10", "MIMS", "mims")),
          starts_with("steps")
        ),
        na_sum
      )
    )
  
  daily <- daily %>%
    full_join(inclusion)
  
  list(
    minute_level = processed,
    inclusion = inclusion,
    daily = daily
  )
}
```

```{r acti_run}
#| cache: true
#| dependson: acti_run_code
out <- acti_run(data)
```

The minute-level output shows the raw time series after step counts, wear-time
classification, and the log-count transform have been added.

```{r minute_level}
out$minute_level %>%
  select(time, counts, starts_with("steps"), counts_log10, date, is_included) %>%
  slice_head(n = 8) %>%
  knitr::kable()
```

The inclusion table records which days met the wear-time threshold. The
threshold is study-specific, but the key idea is that downstream summaries
should only use days with enough valid data.

```{r inclusion}
out$inclusion %>%
  slice_head(n = 8) %>%
  knitr::kable()
```

The daily table is the compact summary that gets paired with the demographic
data.

```{r daily}
out$daily %>%
  slice_head(n = 8) %>%
  knitr::kable()
```

## Reproduce the Demographic Data

The next section reproduces the cleaned demographic table from
[`demog_mapping.R`](./demog_mapping.R). It is wrapped in a collapsible block so
the report stays readable, but the code still runs when the document is rendered.
The chunk is cached and the spreadsheet is stored locally, so the expensive work
only happens once unless the inputs change.

::: {.callout-note collapse="true"}
### Show how the demographic table is rebuilt

```{r}
#| cache: true
if (requireNamespace("readxl", quietly = TRUE)) {
  meta_url <- "https://ndownloader.figshare.com/files/21881742"
  meta_path <- download_if_missing(meta_url, file.path(raw_dir, "Metadata.xlsx"))
  
  meta <- readxl::read_excel(meta_path)
  
  # Excel sometimes creates placeholder columns with names like "..1".
  bad_col <- grepl("^\\.\\.", colnames(meta))
  colnames(meta)[bad_col] <- NA
  
  # The first couple of rows contain extra header information, so we stitch the
  # real column names together before dropping those rows.
  potential_headers <- rbind(colnames(meta), meta[1:2, ])
  potential_headers <- apply(potential_headers, 2, function(x) {
    x <- paste(na.omit(x), collapse = "")
    x <- sub(" .csv", ".csv", x)
    x <- sub(" .wav", ".wav", x)
    x <- gsub(" ", "_", x)
    x
  })
  colnames(meta) <- potential_headers
  
  # The imported table is easiest to use after dropping the extra header rows and
  # normalizing the names to snake_case.
  meta <- meta[-c(1:2), ] %>%
    janitor::clean_names()
  
  # Standardize the participant identifier and clean obvious formatting artifacts.
  meta <- meta %>%
    rename(id = participant_identifier) %>%
    filter(!is.na(id), id != "Participant Identifier") %>%
    mutate(across(where(is.character), ~ gsub("ü", "yes", .x)))
  
  # Convert the fields that should behave like numbers.
  meta <- meta %>%
    mutate(
      age = readr::parse_number(as.character(age)),
      time_since_prescription_of_a_myoelectric_prosthesis_years =
        readr::parse_number(as.character(
          time_since_prescription_of_a_myoelectric_prosthesis_years
        ))
    )
  
  # "Congenital" means there is no time since limb loss, so we encode that as 0.
  meta <- meta %>%
    mutate(
      time_since_limb_loss_years = ifelse(
        time_since_limb_loss_years == "Congenital",
        0,
        time_since_limb_loss_years
      )
    )
  
  demog_lookup <- meta %>%
    filter(id == "AI1") %>%
    select(id, gender, age, right_sensor)
  
  demog_lookup
  
  # The mapping function only needs sex and age, so we convert the cleaned table
  # into the compact input expected by the quantile map.
  demog <- demog_lookup %>%
    transmute(sex = gender, age = age)
  
  demog
  
  saveRDS(demog_lookup, file.path(derived_dir, "demog_lookup_AI1.rds"))
  saveRDS(demog, file.path(derived_dir, "demog_AI1_for_map.rds"))
}
```

:::

## Combine Activity and Demographics

Now we keep only included days, average the daily summaries, and attach the
demographic fields. This is the final data frame passed to the map function.

```{r map_daily}
map_data <- out$daily %>%
  filter(is_included) %>%
  summarise(
    dplyr::across(starts_with("steps"), mean),
    ac = mean(counts)
  ) %>%
  bind_cols(demog) %>%
  pivot_longer(
    cols = c(starts_with("steps"), ac),
    names_to = "measure",
    values_to = "value"
  )

map_data
```

The map itself is a compact visual summary of the mean steps and accelerometer
counts for the included days, positioned relative to the demographic inputs.

```{r map_pa_quantiles}
actiquantiles::acti_map_nhanes(map_data)
```

## Why This Structure Helps

The report separates the analysis into small, named steps so each part is easy
to verify:

1. The raw files are downloaded to persistent paths, so future knits do not
spend time on repeat downloads.
2. The GT3X+ reader works directly on the saved file, which keeps the raw data
untouched.
3. The processing function isolates the transformation logic, making it easier
to test or reuse.
4. The demographic reconstruction is hidden by default but fully documented and
cached.
5. The final mapping step uses the cleaned, joined data rather than a manually
assembled example table.

That makes the document suitable both as a narrative report and as a precise
reproduction of the original analysis workflow.
